---
title: "hands-on-exercise-10"
description: |
  In this lesson, we will be learning about Spatial Interaction Models.
author:
  - name: Toh Jun Long
    url: https://linkedin.com/in/tohjunlong
date: 10-25-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before we getting started, it is important for us to install the necessary R packages and launch them into RStudio environment.

The R packages need for this exercise are as follows:

* Spatial data handling
+ sf, sp, geojsonio, stplanr.
* Attribute data handling
+ tidyverse, especially readr and dplyr, reshape2,
* thematic mapping
+ tmap
* Statistical graphic
+ ggplot2
* Statistical analysis
+ caret

The code chunk below installs and launches these R packages into RStudio environment.

```{r}
packages = c('tmap', 'tidyverse',
             'sp', 'caret',
             'geojsonio', 'reshape2', 
             'broom')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

```{r}
# library(devtools)
# install_version("sf", version = "0.9-8", repos = "http://cran.us.r-project.org")
```

```{r}
# library(devtools)
# install_version("stplanr", version = "0.8.4", repos = "http://cran.us.r-project.org")
```

```{r}
library(sf)
library(stplanr)
```

```{r}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")

```

```{r}
Ausdata <- Aus@data
```

```{r}
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
```

```{r}
st_is_valid(AusSF)
```

```{r}
st_make_valid(AusSF)
```

```{r}
st_is_valid(AusSF)
```

Plot the boundary layer

```{r}
tmap_options(check.and.fix = TRUE)
tmap_mode("plot")
qtm(AusSF)
```

# Displaying data table

```{r}
head(AusSF, 10)
```

```{r}
Aus <- as(AusSF, "Spatial")
```

# Calculating a distance matrix

```{r}
AusProj <- spTransform(Aus,"+init=epsg:3112")
summary(AusProj)
```

```{r}
dist <- spDists(AusProj)
dist 
```

```{r}
distPair <- melt(dist)
head(distPair, 10)
```

```{r}
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

## Converting distance matrix into distance pair list

```{r}
distPair <- melt(dist)
head(distPair, 10)
```

## Converting unit of measurement from metres into km

```{r}
distPair$value <- distPair$value / 1000
head(distPair, 10)

```

# Importing Interaction Data

Importing in the migration data into rstudio

```{r}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)
glimpse(mdata)
```

## Combining the imported migration data

* Add in generated distance data
* Create new column of total flows, excluding flows that occur within an area (intra-area)



```{r}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
```

Ordering the spatial data by code order.

```{r}
mdata$dist <- distPair$value 
```

Setting intr-zonal distance to be near 0 and not exactly 0

```{r}
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
```

```{r}
glimpse(mdata)
```

# Visualising with desire line

We will not plot the intra-zonal flows. The code chunk below will be used to remove intra-zonal flows.

```{r}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]
```

Removing all exceprt origin, dest and flow cols

```{r}
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, 
                          zones = Aus)
```

Converting to WGS84 projection

```{r}
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326" )
```

Same for Aus layer

```{r}
AusWGS <- spTransform(Aus,"+init=epsg:4326" )
```

Setting a more meaningful value for line widths

```{r}
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10
```

Plotting the interaction line onto a map

```{r}
plot(travel_networkwgs, lwd = w)
plot(AusWGS, add=T)
```

# Building Spatial Interaction Models

## Unconstrained Spatial Interaction Model

```{r}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(uncosim)
```

## Fitting the model

```{r}
mdatasub$fitted <- fitted(uncosim)
```

## (Optional) Another method to calculate the estimates.

```{r}
k <- uncosim$coefficients[1]
mu <- uncosim$coefficients[2]
alpha <- uncosim$coefficients[3]
beta <- -uncosim$coefficients[4]
```

```{r}
mdatasub$unconstrainedEst2 <- exp(k+(mu*log(mdatasub$vi1_origpop))+(alpha*log(mdatasub$wj3_destmedinc))-(beta*log(mdatasub$dist)))
```

```{r}
mdatasub$unconstrainedEst2 <- (exp(k)*exp(mu*log(mdatasub$vi1_origpop))*exp(alpha*log(mdatasub$wj3_destmedinc))*exp(-beta*log(mdatasub$dist)))
```

## Saving the fitted values

```{r}
mdatasub$unconstrainedEst2 <- round(mdatasub$unconstrainedEst2,0)
sum(mdatasub$unconstrainedEst2)
```

```{r}
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "unconstrainedEst2", margins=c("Orig_code", "Dest_code"))
mdatasubmat2
```

Compare with original matrix

```{r}
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
mdatasubmat
```

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `unconstrainedEst2`))+
  geom_point(color="black", fill="light blue")
```

## Assessing the model performance

```{r}
postResample(mdatasub$Flow,mdatasub$unconstrainedEst2)
```

# Origin Constrained Spatial Interaction Model

```{r}
origSim <- glm(Flow ~ Orig_code+log(wj3_destmedinc)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
#let's have a look at it's summary...
summary(origSim)
```

```{r}
mdatasub$origSimFitted <- round(fitted(origSim),0)
```

create pivot table to turn paired list into matrix.

```{r}
mdatasubmat3 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "origSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat3
```

```{r}
mdatasubmat
```

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `origSimFitted`))+
  geom_point(color="black", fill="light blue")
```

```{r}
postResample(mdatasub$Flow,mdatasub$origSimFitted)
```

# Destination Constrained Spatial Interaction Model

```{r}
destSim <- glm(Flow ~ Dest_code+log(vi1_origpop)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(destSim)
```

```{r}
mdatasub$destSimFitted <- round(fitted(destSim),0)
```

```{r}
mdatasubmat6 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "destSimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat6
```

```{r}
mdatasubmat
```

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `destSimFitted`))+
  geom_point(color="black", fill="light blue")
```

Test the Goodness-of-Fit in exactly the same way as before:

```{r}
postResample(mdatasub$Flow,mdatasub$destSimFitted)
```

# Doubly Constrained Spatial Interaction Model

Calibrate a Doubly Constrained Spatial Interaction Model by using glm()

```{r}
doubSim <- glm(Flow ~ Orig_code+Dest_code+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)
summary(doubSim)
```

We can examine how the constraints hold for destinations this time. Firstly, we will fitted the model and roundup the estimated values by using the code chunk below.

```{r}
mdatasub$doubsimFitted <- round(fitted(doubSim),0)
```

```{r}
mdatasubmat7 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "doubsimFitted", margins=c("Orig_code", "Dest_code"))
mdatasubmat7
```

Next, we will used the step you had learned in previous section to create pivot table to turn paired list into matrix.

```{r}
mdatasubmat
```

```{r}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `doubsimFitted`))+
  geom_point(color="black", fill="light blue")
```


```{r}
postResample(mdatasub$Flow,mdatasub$doubsimFitted)
```

















