---
title: "Hands-on_Ex4"
description: |
  A short description of the post.
author:
  - name: Toh Jun Long
    url: https://linkedin.com/in/tohjunlong
date: 09-05-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap')
  for (p in packages){
    if(!require(p, character.only = T)){
      install.packages(p)
  }
  library(p,character.only = T)
}
```


```{r}
childcare_sf = st_read("../../Hands-on_Ex04/data/child-care-services-geojson.geojson") %>% 
  st_transform(crs = 3414)
st_geometry(childcare_sf)
```

```{r}
mpsz_sf = st_read(dsn = "../../Hands-on_Ex04/data",
               layer = "MP14_SUBZONE_WEB_PL")
st_geometry(mpsz_sf)
```

```{r}
sg_sf = st_read(dsn = "../../Hands-on_Ex04/data",
               layer = "CostalOutline")
st_geometry(sg_sf)
```
```{r}
st_crs(childcare_sf)
```
```{r}
st_crs(mpsz_sf)
```
```{r}
st_crs(sg_sf)
```
```{r}
mpsz_sf3414 = st_set_crs(mpsz_sf, 3414)
sg_sf3414 = st_set_crs(sg_sf, 3414)
```
```{r}
tm_shape(mpsz_sf3414)+
  tm_polygons()+
tm_shape(childcare_sf)+
  tm_dots()
```
```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```
```{r}
tmap_mode("plot")
```
```{r}
childcare = as_Spatial(childcare_sf)
mpsz = as_Spatial(mpsz_sf3414)
sg = as_Spatial(sg_sf3414)
```
```{r}
childcare
```
```{r}
mpsz
```
```{r}
sg
```
```{r}
childcare_sp = as(childcare,"SpatialPoints")
sg_sp = as(sg,"SpatialPolygons") #why is this polygon and not point, look above
```
```{r}
childcare_sp
```
```{r}
sg_sp
```

Challenge: Do you know what are the differences between Spatial* classes and generic sp object?

Answer: The lack of headers

```{r}
childcare_ppp = as(childcare_sp,"ppp")
childcare_ppp
```
```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```
```{r}
any(duplicated(childcare_ppp))
```
```{r}
multiplicity(childcare_ppp)
```
```{r}
sum(multiplicity(childcare_ppp)>1)
```
```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```
```{r}
tmap_mode("plot")
```
```{r}
childcare_ppp_jit = rjitter(childcare_ppp,
                            retry = TRUE,
                            nsim = 1,
                            drop = TRUE)
```
```{r}
any(duplicated(childcare_ppp_jit))
```
```{r}
sg_owin = as(sg_sp,"owin")
```
```{r}
plot(sg_owin)
```
```{r}
summary(sg_owin)
```
```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
#why not use the one that is jitterred
```
```{r}
summary(childcareSG_ppp)
```
```{r}
plot(childcareSG_ppp)
```
```{r}
kde_childcareSG_bw = density(childcareSG_ppp,
                             sigma = bw.diggle,
                             edge = TRUE,
                             kernel = "gaussian")
```
```{r}
plot(kde_childcareSG_bw)
```

```{r}
bw = bw.diggle(childcareSG_ppp)
bw
```
```{r}
childcareSG_ppp.km = rescale(childcareSG_ppp,1000,"km")
```
```{r}
kde_childcareSG.bw = density(childcareSG_ppp.km,
                             sigma = bw.diggle,
                             edge = TRUE,
                             kernel = "gaussian")
plot(kde_childcareSG.bw)
```
```{r}
bw.CvL(childcareSG_ppp.km)
```
```{r}
bw.scott(childcareSG_ppp.km)
```
```{r}
bw.ppl(childcareSG_ppp.km)
```
```{r}
bw.diggle(childcareSG_ppp.km)
```
```{r}
kde_childcareSG.ppl = density(childcareSG_ppp.km,
                              sigma = bw.ppl,
                              edge = TRUE,
                              kernel = "gaussian")
par(mfrow = c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```
```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```
```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km,
                               sigma=0.6,
                               edge=TRUE,
                               kernel="gaussian")
plot(kde_childcareSG_600)
```
```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km,                                      method="kernel")
plot(kde_childcareSG_adaptive)
```
```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```
```{r}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```
```{r}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```
```{r}
kde_childcareSG_bw_raster

```
```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```
```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```
```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```
```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```
```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```
```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```
```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```
```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```
```{r}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```
```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```
```{r}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

based on clark-evans test, the p-value is smaller than alpha of 0.5%, We should this reject the null-hypothesis that the point pattern are randomly distributed

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)

```

Similarly, the spatial distribution of the spatial points within the choa chu kang region is suggested to be non-random as the p-value is smaller than the alpha of 0.5.

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

Similarly, the spatial distribution of the spatial points within the Tampines region is suggested to be non-random as the p-value is smaller than the alpha of 0.5.

Second order analysis:

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```
```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```
```{r}
plot(G_CK.csr)
```

Statistically insignificant

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
plot(G_tm.csr)
```

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```
Hypo testing for F function for cck

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```
```{r}
plot(F_CK.csr)
```

Statistically insignificant

Tampines (F-function):

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

Hypo testing:

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

2nd order effect spatial point process analysis using K-function

Choa chu kang:

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

Hypo testing:

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr)
```


Tampines:

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

Hypo testing:

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr)
```

2nd order effect spatial point process analysis using L-function


Choa chu kang:

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

Hypo testing with 99 sim:

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

Tampines:

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

hypo testing with 99 sim:

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

normalised around 0:

```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```






