---
title: "in-class-exercise-4"
description: |
  A short description of the post.
author:
  - name: Toh Jun Long
    url: https://linkedin.com/in/tohjunlong
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
#get_theme(one_map_token,"")
```

- sf is mainly used for importing geospatial data

- spatstat is used to perform spatial point patern analysis, might be a little outdated

- maptools is used for map data conversion

- ppp is the format supported by spatstat, hence we need maptools to convert spatial data from one type to ppp format

- raster is used for handling raster-based geospatial data

- tmap for mapping

- tidyverse for non-spatial data

```{r echo = TRUE}
packages = c("maptools","sf","raster","spatstat","tmap","tidyverse")
 for (p in packages){
    if(!require(p, character.only = T)){
      install.packages(p)
  }
  library(p,character.only = T)
}
```
```{r echo = TRUE}
sg_sf = st_read(dsn = "data/shapefile",
                layer = "CostalOutline")
mpsz_sf = st_read(dsn = "data/shapefile",
                layer ="MP14_SUBZONE_WEB_PL")
```

This is a new read format,using rds file as input file type. 

Childcare layer is just attribute table, aspatial data, no geometric data. Likewise for chas data.

We need to check the data even if the dataset come from the same agency, structure of a dataset may vary.


```{r echo = TRUE}
childcare = read_rds("data/rds/childcare.rds")
chas = read_rds("data/rds/CHAS.rds")
```

### Converting from aspatial to geospatial

```{r echo = TRUE}
chas_sf = st_as_sf(chas, coords = c("X_COORDINATE","Y_COORDINATE"),
                   crs = 3414)
```

4326 is the EPSG for WGS84, so we need to set it first then convert the coordinate system.

SVY21 is in metres while WGS84 is in degrees

data scientist at data.gov usually handle geospatial data using kml or geojson file type

thus the API is introduced for a better structured geospatial data format

As a data analyst, we should choose the data that can benefit us the most when performing analysis

```{r echo = TRUE}
childcare$Lat = as.numeric(childcare$Lat)
childcare$Lng = as.numeric(childcare$Lng)
childcare_sf = st_as_sf(childcare, 
                        coords = c("Lng",
                                   "Lat"),
                        crs = 4326) %>% 
  st_transform(crs = 3414)
```

### converting from sf format to Spatial* classes format

```{r echo = TRUE}
childcare = as_Spatial(childcare_sf)
chas = as_Spatial(chas_sf)
mpsz = as_Spatial(mpsz_sf)
sg = as_Spatial(sg_sf)
```

### converting from Spatial* classes to sp format

SP layer do not have attribute/associate table

```{r echo = TRUE}
childcare_sp = as(childcare, "SpatialPoints")
chas_sp = as(chas, "SpatialPoints")
sg_sp = as(sg, "SpatialPolygons")
```

### converting from sp to spatstat ppp format

Provides owin, the window study area, would be rectangle based on the xlim and ylim of the data

once converted, projection would be lost, when converting back to raster layer, we need to add in the projection

```{r echo = TRUE}
childcare_ppp = as(childcare_sp,"ppp")
chas_ppp = as(chas_sp,"ppp")
```

Keep interactive maps below 10 maps, we need to be aware of the limits web servers have

```{r echo = TRUE}
tmap_mode("view")
tm_shape(childcare_sf)+
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05) + 
tm_shape(chas_sf) + 
  tm_dots(alpha = 0.4,
          col = "red",
          size = 0.05)
```

converting tmap_mode back to plot

```{r echo = TRUE}
tmap_mode("plot")
```





