---
title: "Take Home Exercise 1"
description: |
  A short description of the post.
author:
  - name: Toh Jun Long
    url: https://linkedin.com/in/tohjunlong
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

1. Introduction

Since late December 2019, an outbreak of a novel coronavirus disease (COVID-19; previously known as 2019-nCoV) was reported in Wuhan, China, which had subsequently affected 210 countries worldwide. In general, COVID-19 is an acute resolved disease but it can also be deadly, with a 2% case fatality rate.

The COVID-19 pandemic in Indonesia is part of the ongoing worldwide pandemic of coronavirus disease 2019 (COVID-19) caused by severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2). The virus was confirmed to have reached Indonesia on March 2, 2020. It started with two cases in March. As of July 31 2021, there had been 3,409,658 cumulative confirmed cases of COVID-19 in Indonesia and 94,119 reported cumulative deaths. All cases were spread in 34 provinces in Indonesia. Among all the provinces, DKI Jakarta (Indonesian: Daerah Khusus Ibukota Jakarta and in English: Special Capital Region of Jakarta) contributed close to 24% of the cumulative confirmed cases.

Despite its compactness, the cumulative confirmed cases were not evenly distributed within DKI Jakarta as Figure below reveals at the district and sub-district levels. The question was where were the sub-districts with relatively higher number of confirmed cases and how they changed over time.

Goals:
- Identify the sub-districts with higher number of confirmed cases than average
- Visualize the change in COVID-19 infection from March 2020 to August 2021

2. Data

Data: 
  1. Open Data Covid-19 Provinsi DKI Jakarta
    - source: Riwayat File Covid-19 DKI Jakarta
    - Format: xlsx
  2. Shapefile (SHP) Batas Desa Provinsi DKI Jakarta
    - source: PODES 2019
    - Format: shp


3. Installing and loading the necessary packages if not installed

Packages used: maptools, sf, raster, spatstat and tmap, tidyverse, readxl

```{r echo=TRUE, message=FALSE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap','tidyverse','readxl')
for(p in packages){
  if(!require(p,character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

4. Importing the data

Reading the datas

```{r echo=TRUE}
jakarta = st_read(dsn = "data/Geospatial",
                  layer = "BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")
```

Check crs

```{r echo=TRUE}
st_crs(jakarta)
```

Upon Reading the CRS, we can see that the projected coordinate systrem is different from the designated system of Indoneisa, which is DGN95 / Indonesia TM-3 zone 54.1, EPSG: 23845.

We then have to transform the projection to EPSG:23845.

```{r echo=TRUE}
jakarta_sf23845 = st_transform(jakarta, 23845)
st_crs(jakarta_sf23845)
```

The change has been completed. WE can use st_geometry to take a look at the sf attributes.

```{r echo=TRUE}
st_geometry(jakarta_sf23845)
```


Finding the common values of outer islands. By clicking at a few of the outer islands, we can see that one common data value is "KEPULAUAN SERIBU" under "KAB_KOTA"(City District) attribute.

```{r echo=TRUE}
tmap_mode("view")
tm_shape(jakarta_sf23845)+
  tm_polygons()
```

We can add in colors to confirm that the outer islands have a common value for KAB_KOTA

```{r echo=TRUE}
tm_shape(jakarta_sf23845)+
  tm_polygons() +
tm_shape(jakarta_sf23845) +
  tm_fill("KAB_KOTA",
          palette = "RdYlBu")+
  tm_borders()
```

setting tmap_mode back to plot

```{r echo=TRUE}
tmap_mode("plot")
```

Using the colSums to check the number of missing values in each attribute. From this we can see that there are at most 2 unique missing value. We need to check if we can remove the 2 value or we should keep it.

```{r echo = TRUE}
colSums(is.na(jakarta_sf23845))
```

Using rowSums and extracting any rows with missing value, we can see that it is truly 2 unique row that contains missing values in some of the columns, however, the data can still be kept within the dataframe as it provides geospatial properties that is part of a full constructed map.

```{r echo=TRUE}
jakarta_sf23845[rowSums(is.na(jakarta_sf23845)) >0,]
```

Instead of dropping the rows with missing values, we can fill in with a string "Missing" as a placeholder. This can only be done if the columns do not contain important information that is useful for further processing

```{r echo=TRUE, message=FALSE}
jakarta_sf23845[is.na(jakarta_sf23845)] = "Missing"
```

Dropping the rows that has "KEPULAUAN SERIBU" in column, KAB_KOTA

```{r echo=TRUE, message=FALSE}
jakarta_sf23845_cleaned = jakarta_sf23845[jakarta_sf23845$KAB_KOTA != "KEPULAUAN SERIBU",]
```

checking geometry validity

```{r echo=TRUE}
length(which(st_is_valid(jakarta_sf23845_cleaned) == FALSE))
```

Confirming that external islands are removed

```{r echo=TRUE, message=FALSE}
tm_shape(jakarta_sf23845_cleaned)+
  tm_polygons()
```

Retrieving first 9 columns of the data frame to be the cleaned data frame

```{r echo=TRUE}
jakarta_sf23845_cleaned = jakarta_sf23845_cleaned[,c(1:9)]
head(jakarta_sf23845_cleaned,5)
```

Importing Aspatial data (Daily COVID 19 DKI Jakarta)

```{r echo=TRUE}
jakarta_covid = read_excel("data/Aspatial/Daily COVID 19 DKI Jakarta.xlsx") 
jakarta_covid = jakarta_covid[-1,]
list(jakarta_covid)
```

Geospatial Data Integration:

Creating a function to initialise month data frame.

```{r echo=TRUE, message=FALSE}
init_month = function(df, month){
  result = df[df$Month_year == month,]
  return(result)
}
```

Creating a function to perform left_join from aspatial dataframe to geospatial dataframe.

```{r echo=TRUE, message=FALSE}
join_funct = function(aspatial_df){
  result = left_join(jakarta_sf23845_cleaned,aspatial_df,by = c("DESA_KELUR" = "nama_kelurahan")) %>% 
    mutate(`cumulative confirmed cases rate` = (`POSITIF`/`JUMLAH_PEN`)) %>% 
    mutate(`cumulative confirmed death rate` = (`Meninggal`/`JUMLAH_PEN`))
  return(result)
}
```

using both function to initialize and join to create covid-19 dataframe of each month

```{r echo=TRUE, message=FALSE}
mar_20_covid = join_funct(init_month(jakarta_covid,"Mar-20"))
apr_20_covid = join_funct(init_month(jakarta_covid,"Apr-20"))
may_20_covid = join_funct(init_month(jakarta_covid,"May-20"))
jun_20_covid = join_funct(init_month(jakarta_covid,"Jun-20"))
jul_20_covid = join_funct(init_month(jakarta_covid,"Jul-20"))
aug_20_covid = join_funct(init_month(jakarta_covid,"Aug-20"))
sep_20_covid = join_funct(init_month(jakarta_covid,"Sep-20"))
oct_20_covid = join_funct(init_month(jakarta_covid,"Oct-20"))
nov_20_covid = join_funct(init_month(jakarta_covid,"Nov-20"))
dec_20_covid = join_funct(init_month(jakarta_covid,"Dec-20"))
jan_21_covid = join_funct(init_month(jakarta_covid,"Jan-21"))
feb_21_covid = join_funct(init_month(jakarta_covid,"Feb-21"))
mar_21_covid = join_funct(init_month(jakarta_covid,"Mar-21"))
apr_21_covid = join_funct(init_month(jakarta_covid,"Apr-21"))
may_21_covid = join_funct(init_month(jakarta_covid,"May-21"))
jun_21_covid = join_funct(init_month(jakarta_covid,"Jun-21"))
jul_21_covid = join_funct(init_month(jakarta_covid,"Jul-21"))
aug_21_covid = join_funct(init_month(jakarta_covid,"Aug-21"))
```

6. Maps

6.1 Thematic Mapping

6.1.1 At a glance:

Distribution of COVID-19 cumulative positive and death cases up till 31 December 2020

```{r echo=TRUE, message=FALSE}
tm_shape(jakarta_sf23845_cleaned)+
  tm_polygons()+
tm_shape(dec_20_covid) +
  tm_fill(c("POSITIF","Meninggal"),
          style = "jenks",
          palette = "Blues")+
  tm_layout(legend.outside = TRUE,
            legend.position = c("right","bottom"))+
  tm_borders()
```

From this we can see that some areas have both high postive cases and death rates for the daily data until 31 dec 2020.

Distribution of COVID-19 cumulative positive and death cases up till 31 August 2021

```{r echo=TRUE, message=FALSE}
tm_shape(jakarta_sf23845_cleaned)+
  tm_polygons()+
tm_shape(aug_21_covid) +
  tm_fill(c("POSITIF","Meninggal"),
          style = "jenks",
          palette = "Blues")+
  tm_layout(legend.outside = TRUE,
            legend.position = c("right","bottom"))+
  tm_borders()
```

From this we can see that similar to dec 2020, some areas have both high postive cases and death rates for the daily data until 31 Aug 2021, however, the number of cases also increased from max of 1360 to 6547.


Function for creating standard map with Quanitle distribution

```{r echo=TRUE, message=FALSE}
map1 = function(df, vname, legtitle = NA,mtitle = NA){
  tm_shape(df) +
    tm_fill(vname,
            style = "quantile",
            palette = "Blues",
            title = legtitle
            )+
    tm_layout(main.title = mtitle,
              main.title.position = "center",
              main.title.size = 0.8,
              legend.outside = TRUE,
              legend.position = c("right","bottom"),
              frame = TRUE
              )+
    tm_borders(alpha = 0.5)+
    tm_scale_bar(width = 0.15) +
    tm_grid(lwd = 0.1, alpha = 0.2) +
    tm_credits("Source: Open Data Covid-19 Provinsi DKI Jakarta. from Riwayat File Covid-19 DKI Jakarta", 
               position = c("left", "bottom"))
}

```

Creating the individual cumulative confirmed cases maps for each month.

```{r echo=TRUE, message=FALSE, warning=FALSE}
cumu_case_mar_20 = map1(mar_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","March 2020")
cumu_case_apr_20 = map1(apr_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","April 2020")
cumu_case_may_20 = map1(may_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","May 2020")
cumu_case_jun_20 = map1(jun_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","June 2020")
cumu_case_jul_20 = map1(jul_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","July 2020")
cumu_case_aug_20 = map1(aug_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","August 2020")
cumu_case_sep_20 = map1(sep_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","September 2020")
cumu_case_oct_20 = map1(oct_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","October 2020")
cumu_case_nov_20 = map1(nov_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","November 2020")
cumu_case_dec_20 = map1(dec_20_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","December 2020")
cumu_case_jan_21 = map1(jan_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","January 2021")
cumu_case_feb_21 = map1(feb_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","February 2021")
cumu_case_mar_21 = map1(mar_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","March 2021")
cumu_case_apr_21 = map1(apr_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","April 2021")
cumu_case_may_21 = map1(may_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","May 2021")
cumu_case_jun_21 = map1(jun_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","June 2021")
cumu_case_jul_21 = map1(jul_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","July 2021")
cumu_case_aug_21 = map1(aug_21_covid,"cumulative confirmed cases rate","Cumulative Postive Cases","August 2021")

```

Using tmap_arrange to arrange the individual map to into a 4 (columns) by 3 (rows) layout.

Cumulative Confirmed Cases rate for each month from March 2020 to June 2020 (4 Months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_case_mar_20,cumu_case_apr_20,cumu_case_may_20,cumu_case_jun_20, asp = 1, ncol = 2)
```

Cumulative Confirmed Cases rate for each month from July 2020 to October 2020 (4 Months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_case_jul_20,cumu_case_aug_20,cumu_case_sep_20,cumu_case_oct_20, asp = 1, ncol = 2)
```

Cumulative Confirmed Cases rate for each month from November 2020 to February 2021 (4 Months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_case_nov_20,cumu_case_dec_20,cumu_case_jan_21,cumu_case_feb_21, asp = 1, ncol = 2)
```

Cumulative Confirmed Cases rate for each month from March 2021 to June 2021 (4 Months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_case_mar_21,cumu_case_apr_21,cumu_case_may_21,cumu_case_jun_21, asp = 1, ncol = 2)
```

Cumulative Confirmed Cases rate for each month from July 2021 to August 2021 (2 Months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_case_jul_21,cumu_case_aug_21, asp = 1, ncol = 2)
```

From the graphical visualization, we can see that the while the cumulative positive case rate increases to 0.394 confirmed case per population, the spread of the confirmed cases starts more severely in north Jakarta in March 2020, before moving towards south Jakarta by Aug 2021

Creating the individual cumulative death cases maps for each month.

```{r echo=TRUE, message=FALSE, warning = FALSE}
cumu_death_mar_20 = map1(mar_20_covid,"cumulative confirmed death rate","Death Cases","March 2020")
cumu_death_apr_20 = map1(apr_20_covid,"cumulative confirmed death rate","Death Cases","April 2020")
cumu_death_may_20 = map1(may_20_covid,"cumulative confirmed death rate","Death Cases","May 2020")
cumu_death_jun_20 = map1(jun_20_covid,"cumulative confirmed death rate","Death Cases","June 2020")
cumu_death_jul_20 = map1(jul_20_covid,"cumulative confirmed death rate","Death Cases","July 2020")
cumu_death_aug_20 = map1(aug_20_covid,"cumulative confirmed death rate","Death Cases","August 2020")
cumu_death_sep_20 = map1(sep_20_covid,"cumulative confirmed death rate","Death Cases","September 2020")
cumu_death_oct_20 = map1(oct_20_covid,"cumulative confirmed death rate","Death Cases","October 2020")
cumu_death_nov_20 = map1(nov_20_covid,"cumulative confirmed death rate","Death Cases","November 2020")
cumu_death_dec_20 = map1(dec_20_covid,"cumulative confirmed death rate","Death Cases","December 2020")
cumu_death_jan_21 = map1(jan_21_covid,"cumulative confirmed death rate","Death Cases","January 2020")
cumu_death_feb_21 = map1(feb_21_covid,"cumulative confirmed death rate","Death Cases","February 2020")
cumu_death_mar_21 = map1(mar_21_covid,"cumulative confirmed death rate","Death Cases","March 2020")
cumu_death_apr_21 = map1(apr_21_covid,"cumulative confirmed death rate","Death Cases","April 2020")
cumu_death_may_21 = map1(may_21_covid,"cumulative confirmed death rate","Death Cases","May 2020")
cumu_death_jun_21 = map1(jun_21_covid,"cumulative confirmed death rate","Death Cases","June 2020")
cumu_death_jul_21 = map1(jul_21_covid,"cumulative confirmed death rate","Death Cases","July 2020")
cumu_death_aug_21 = map1(aug_21_covid,"cumulative confirmed death rate","Death Cases","August 2020")
```

Using tmap_arrange to arrange the individual map to into a 4 (columns) by 3 (rows) layout.

Cumulative death rate for each month from March 2020 to June 2020 (4 months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_death_mar_20,cumu_death_apr_20,cumu_death_may_20,cumu_death_jun_20, asp = 1, ncol = 2)
```

Cumulative death rate for each month from July 2020 to October 2020 (4 months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_death_jul_20,cumu_death_aug_20,cumu_death_sep_20,cumu_death_oct_20, asp = 1, ncol = 2)
```

Cumulative death rate for each month from November 2020 to February 2021 (4 months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_death_nov_20,cumu_death_dec_20,cumu_death_jan_21,cumu_death_feb_21, asp = 1, ncol = 2)
```

Cumulative death rate for each month from March 2020 to June 2021 (4 months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_death_mar_21,cumu_death_apr_21,cumu_death_may_21,cumu_death_jun_21, asp = 1, ncol = 2)
```

Cumulative death rate for each month from July 2021 to August 2021 (2 months)

```{r echo=TRUE, message=FALSE, warning = FALSE}
tmap_arrange(cumu_death_jul_21,cumu_death_aug_21, asp = 1, ncol = 2)

```

From the graphical visualization, we can see that the while the cumulative death rate increases to 0.0042 death toll per population, most of the highest death rate occurs around the central of Jakarta.


